---
title: "Thesis Analysis"
output: 
  pdf_document:
    keep_tex: yes
date: "2023-02-22"
---


```{r setup, include=FALSE}
#test
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
setwd("~/senior project")
library(readxl)
library(tidyverse)
library(lubridate)
library(Hmisc)
library(heplots)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(caret)
library(class)
library(kableExtra)
library(knitr)
library(car)
library(ggeffects)
#library(interplot)
#test2
```

# Secondary Cleaning
## loading data
```{r}
# # flow >> rename geographies, assign mutual geos to payscale data and census 
# >> filter race data to groups we care about >> add entropy values >> join 
# census race data with payscale



### Filter payscale data by race
# loading data
load("~/senior project/alldata.RData") # original dataset

# subset based on 5 major race groups recognized by census
# organize race
head(sort(table(alldata$RACE_SET),TRUE))
alldata <- alldata %>% 
  filter(RACE_SET %in% c("White", "Hispanic", "Black or African American",
                         "Asian", "American Indian and Alaska Native"))
table(alldata$RACE_SET)
```
```{r}
### read in census datasets
# note cbsa_race10.csv in my filesystem is table P1, v2 is table P2 
# from census bureau
# cbsa_race10 is table p2 from 2010
cbsa_race10 <- read.csv("cbsa_race10v2.csv")
# cbsa_city is a reference table of major cities within each CBSA
cbsa_city <- read_xls("cbsa_major_cities.xls", col_names = TRUE)
cbsa_city <- cbsa_city[-c(1:2),] # gets rid of extra header
names(cbsa_city) <- c("CBSA_code", "CBSA_title", "metro_micro_area", 
                      "principal_city", "FIPS_state", "FIPS_place")
```

## renaming locations 
```{r}
### assigning a CBSA to each datapoint in alldata
#recoding states in payscale data
alldata <- alldata %>% 
  mutate(
    LOCATION_STATE = case_when(
      LOCATION_STATE == "Alabama" ~ 'AL',
      LOCATION_STATE == "Alaska" ~ 'AK',
      LOCATION_STATE == "Arizona" ~ 'AZ',
      LOCATION_STATE == "Arkansas" ~ 'AR',
      LOCATION_STATE == "California" ~ 'CA',
      LOCATION_STATE == "Colorado" ~ 'CO',
      LOCATION_STATE == "Connecticut" ~ 'CT',
      LOCATION_STATE == "Delaware" ~ 'DE',
      LOCATION_STATE == "Florida" ~ 'FL',
      LOCATION_STATE == "Georgia" ~ 'GA',
      LOCATION_STATE == "Hawaii" ~ 'HI',
      LOCATION_STATE == "Idaho" ~ 'ID',
      LOCATION_STATE == "Illinois" ~ 'IL',
      LOCATION_STATE == "Indiana" ~ 'IN',
      LOCATION_STATE == "Iowa" ~ 'IA',
      LOCATION_STATE == "Kansas" ~ 'KS',
      LOCATION_STATE == "Kentucky" ~ 'KY',
      LOCATION_STATE == "Louisiana" ~ 'LA',
      LOCATION_STATE == "Maine" ~ 'ME',
      LOCATION_STATE == "Maryland" ~ 'MD',
      LOCATION_STATE == "Massachusetts" ~ 'MA',
      LOCATION_STATE == "Michigan" ~ 'MI',
      LOCATION_STATE == "Minnesota" ~ 'MN',
      LOCATION_STATE == "Mississippi" ~ 'MS',
      LOCATION_STATE == "Missouri" ~ 'MO',
      LOCATION_STATE == "Montana" ~ 'MT',
      LOCATION_STATE == "Nebraska" ~ 'NE',
      LOCATION_STATE == "Nevada" ~ 'NV',
      LOCATION_STATE == "New Hampshire" ~ 'NH',
      LOCATION_STATE == "New Jersey" ~ 'NJ',
      LOCATION_STATE == "New Mexico" ~ 'NM',
      LOCATION_STATE == "New York" ~ 'NY',
      LOCATION_STATE == "North Carolina" ~ 'NC',
      LOCATION_STATE == "North Dakota" ~ 'ND',
      LOCATION_STATE == "Ohio" ~ 'OH',
      LOCATION_STATE == "Oklahoma" ~ 'OK',
      LOCATION_STATE == "Oregon" ~ 'OR',
      LOCATION_STATE == "Pennsylvania" ~ 'PA',
      LOCATION_STATE == "Rhode Island" ~ 'RI',
      LOCATION_STATE == "South Carolina" ~ 'SC',
      LOCATION_STATE == "South Dakota" ~ 'SD',
      LOCATION_STATE == "Tennessee" ~ 'TN',
      LOCATION_STATE == "Texas" ~ 'TX',
      LOCATION_STATE == "Utah" ~ 'UT',
      LOCATION_STATE == "Vermont" ~ 'VT',
      LOCATION_STATE == "Virginia" ~ 'VA',
      LOCATION_STATE == "Washington" ~ 'WA',
      LOCATION_STATE == "West Virginia" ~ 'WV',
      LOCATION_STATE == "Wisconsin" ~ 'WI',
      LOCATION_STATE == "Wyoming" ~ 'WY',
      LOCATION_STATE == "District of Columbia" ~ 'DC',
      TRUE ~ "NA"
    )
  )

# recoding states in cbsa_city (reference table) to match LOCATION_STATE in 
# payscale data
# needs to be numeric for case_when function
cbsa_city$FIPS_state <- as.numeric(cbsa_city$FIPS_state)
cbsa_city <- cbsa_city %>%
  mutate(
    LOCATION_STATE = case_when(
      FIPS_state == 01 ~ "ALABAMA",
      FIPS_state == 02 ~ "ALASKA",
      FIPS_state == 04 ~ "ARIZONA",
      FIPS_state == 05 ~ "ARKANSAS",
      FIPS_state == 06 ~ "CALIFORNIA",
      FIPS_state == 08 ~ "COLORADO",
      FIPS_state == 09 ~ "CONNECTICUT",
      FIPS_state == 10 ~ "DELAWARE",
      FIPS_state == 11 ~ "DISTRICT OF COLUMBIA",
      FIPS_state == 12 ~ "FLORIDA",
      FIPS_state == 13 ~ "GEORGIA",
      FIPS_state == 15 ~ "HAWAII",
      FIPS_state == 16 ~ "IDAHO",
      FIPS_state == 17 ~ "ILLINOIS",
      FIPS_state == 18 ~ "INDIANA",
      FIPS_state == 19 ~ "IOWA",
      FIPS_state == 20 ~ "KANSAS",
      FIPS_state == 21 ~ "KENTUCKY",
      FIPS_state == 22 ~ "LOUISIANA",
      FIPS_state == 23 ~ "MAINE",
      FIPS_state == 24 ~ "MARYLAND",
      FIPS_state == 25 ~ "MASSACHUSETTS",
      FIPS_state == 26 ~ "MICHIGAN",
      FIPS_state == 27 ~ "MINNESOTA",
      FIPS_state == 28 ~ "MISSISSIPPI",
      FIPS_state == 29 ~ "MISSOURI",
      FIPS_state == 30 ~ "MONTANA",
      FIPS_state == 31 ~ "NEBRASKA",
      FIPS_state == 32 ~ "NEVADA",
      FIPS_state == 33 ~ "NEW HAMPSHIRE",
      FIPS_state == 34 ~ "NEW JERSEY",
      FIPS_state == 35 ~ "NEW MEXICO",
      FIPS_state == 36 ~ "NEW YORK",
      FIPS_state == 37 ~ "NORTH CAROLINA",
      FIPS_state == 38 ~ "NORTH DAKOTA",
      FIPS_state == 39 ~ "OHIO",
      FIPS_state == 40 ~ "OKLAHOMA",
      FIPS_state == 41 ~ "OREGON",
      FIPS_state == 42 ~ "PENNSYLVANIA",
      FIPS_state == 44 ~ "RHODE ISLAND",
      FIPS_state == 45 ~ "SOUTH CAROLINA",
      FIPS_state == 46 ~ "SOUTH DAKOTA",
      FIPS_state == 47 ~ "TENNESSEE",
      FIPS_state == 48 ~ "TEXAS",
      FIPS_state == 49 ~ "UTAH",
      FIPS_state == 50 ~ "VERMONT",
      FIPS_state == 51 ~ "VIRGINIA",
      FIPS_state == 53 ~ "WASHINGTON",
      FIPS_state == 54 ~ "WEST VIRGINIA",
      FIPS_state == 55 ~ "WISCONSIN",
      FIPS_state == 56 ~ "WYOMING",
      TRUE ~ "NA"
    )
  )
# shorten to abbrev, could prob do more elegantly but eh
cbsa_city <- cbsa_city %>% 
  mutate(
    LOCATION_STATE = case_when(
      LOCATION_STATE == "ALABAMA" ~ "AL",
      LOCATION_STATE == "ALASKA" ~ "AK",
      LOCATION_STATE == "ARIZONA" ~ "AZ",
      LOCATION_STATE == "ARKANSAS" ~ "AR",
      LOCATION_STATE == "CALIFORNIA" ~ "CA",
      LOCATION_STATE == "COLORADO" ~ "CO",
      LOCATION_STATE == "CONNECTICUT" ~ "CT",
      LOCATION_STATE == "DELAWARE" ~ "DE",
      LOCATION_STATE == "DISTRICT OF COLUMBIA" ~ "DC",
      LOCATION_STATE == "FLORIDA" ~"FL",
      LOCATION_STATE == "GEORGIA" ~ "GA",
      LOCATION_STATE == "HAWAII" ~ "HI",
      LOCATION_STATE == "IDAHO" ~ "ID",
      LOCATION_STATE == "ILLINOIS" ~ "IL",
      LOCATION_STATE == "INDIANA" ~ "IN",
      LOCATION_STATE == "IOWA" ~ "IA",
      LOCATION_STATE == "KANSAS" ~ "KS",
      LOCATION_STATE == "KENTUCKY" ~ "KY",
      LOCATION_STATE == "LOUISIANA" ~ "LA",
      LOCATION_STATE == "MAINE" ~ "ME",
      LOCATION_STATE == "MARYLAND" ~ "MD",
      LOCATION_STATE == "MASSACHUSETTS" ~ "MA",
      LOCATION_STATE == "MICHIGAN" ~ "MI",
      LOCATION_STATE == "MINNESOTA" ~ "MN",
      LOCATION_STATE == "MISSISSIPPI" ~ "MS",
      LOCATION_STATE == "MISSOURI" ~ "MO",
      LOCATION_STATE == "MONTANA" ~ "MT",
      LOCATION_STATE == "NEBRASKA" ~ "NE",
      LOCATION_STATE == "NEVADA" ~ "NV",
      LOCATION_STATE == "NEW HAMPSHIRE" ~ "NH",
      LOCATION_STATE == "NEW JERSEY" ~ "NJ",
      LOCATION_STATE == "NEW MEXICO" ~ "NM",
      LOCATION_STATE == "NEW YORK" ~ "NY",
      LOCATION_STATE == "NORTH CAROLINA" ~ "NC",
      LOCATION_STATE == "NORTH DAKOTA" ~ "ND",
      LOCATION_STATE == "OHIO" ~ "OH",
      LOCATION_STATE == "OKLAHOMA" ~ "OK",
      LOCATION_STATE == "OREGON" ~ "OR",
      LOCATION_STATE == "PENNSYLVANIA" ~ "PA",
      LOCATION_STATE == "RHODE ISLAND" ~ "RI",
      LOCATION_STATE == "SOUTH CAROLINA" ~ "SC",
      LOCATION_STATE == "SOUTH DAKOTA" ~ "SD",
      LOCATION_STATE == "TENNESSEE" ~ "TN",
      LOCATION_STATE == "TEXAS" ~ "TX",
      LOCATION_STATE == "UTAH" ~ "UT",
      LOCATION_STATE == "VERMONT" ~ "VT",
      LOCATION_STATE == "VIRGINIA" ~ "VA",
      LOCATION_STATE == "WASHINGTON" ~ "WA",
      LOCATION_STATE == "WEST VIRGINIA" ~ "WV",
      LOCATION_STATE == "WISCONSIN" ~ "WI",
      LOCATION_STATE == "WYOMING" ~ "WY",
      TRUE ~ "NA"
    )
  )

# getting consistent city, state tags for payscale data
alldata$citystate <- stringr::str_c(alldata$LOCATION_CITY, ", ", 
                                    alldata$LOCATION_STATE)
# func to get state from cbsa titles
# substrRight <- function(x, n){
#   substr(x, nchar(x)-n+1, nchar(x))
# }


# gettting consistent city, state tags for cbsa reference table
cbsa_city$citystate <- stringr::str_c(cbsa_city$principal_city, ", ",
                                      cbsa_city$LOCATION_STATE)

# now cbsa_city and alldata have a column I can use to join in official
# metro/micro areas
```





```{r}
# adding metro areas to the payscale data via left join
# only 56 datapoints don't fit into a metro/micro area
cbsa_city_small <- cbsa_city[,c("CBSA_title", "citystate", "metro_micro_area")]
alldata <- alldata %>% left_join(cbsa_city_small, by = "citystate")
# alldata[,c("citystate", "LOCATION_CITY", "LOCATION_STATE")]
```

```{r}
# adding race data from table P2 from the decennial census to payscale data
# excluding data on specific multiracial groups
cbsa_race10 <- cbsa_race10[complete.cases(cbsa_race10),c(1:12)]
# converting to numeric
for (i in c(2:12)){
  cbsa_race10[,i] <- as.numeric(gsub(",", "", cbsa_race10[,i]))
}
# renaming census columns to be joined with payscale data
names(cbsa_race10) <- c("CBSA_title", "t_2010", "t_hispanic_2010",
                        "t_not_hisp_2010", "t_single_2010", "t_white_2010", 
                        "t_black_2010", "t_native_2010", "t_asian_2010",
                        "t_islander_2010", "other_2010", "multi_2010")

```


## Doing it all again for 2020 lol
```{r}
cbsa_race20 <- read.csv("cbsa_race20v2.csv")
cbsa_race20 <- cbsa_race20[complete.cases(cbsa_race20),c(1:12)]
# converting to numeric
for (i in c(2:12)){
  cbsa_race20[,i] <- as.numeric(gsub(",", "", cbsa_race20[,i]))
}
# renaming census columns to be joined with payscale data
names(cbsa_race20) <- c("CBSA_title", "t_2020", "t_hispanic_2020",
                        "t_not_hisp_2020", "t_single_2020", "t_white_2020", 
                        "t_black_2020", "t_native_2020", "t_asian_2020",
                        "t_islander_2020", "other_2020", "multi_2020")

```


## creating entropy rate

Recall Shannon's definition of information entropy: 
$$ \sum p log(\frac{1}{p}) $$
where p is the proportion of a given group in the overall population.

I will create this variable for the cbsa_raceXX datasets before merging them
with the larger payScale dataset. This should be much less computationally 
expensive.

```{r}
#####################################
# function to calculate entropy
#####################################

# first calculate entropy of each cbsa area (do this before the join!!!)
# we will use the cbbsa_city_race10 and cbsa_race20 tables
# recall main groups are: hispanic, white, black, native american, asian, pacific islander, other single race, and multiracial

# this function will take an individual row from cbsa_raceXX, and calculate the
# entropy rate for that city
# we can iterate this function over all rows of cbsa_raceXX so when we join,
# each observation will have an associated entropy for 2010 and 2020
entropyrate <- function(x){
  racecounts <- as.vector(x[c(3,6,7,8,9,10,11,12)])
  props <- as.numeric(racecounts) / as.numeric(x[2])
  entropies <- props * log(1/props)
  entropy <- sum(entropies)
  return(entropy)
} ############## LOL PLEASE CHECK THIS

### seems right!
# entropyrate(cbsa_race10[1,])
```

```{r}
###############################
# applying entropy function to cbsa data
###############################
cbsa_race10 <- cbsa_race10 %>% 
  mutate(
    entropy10 = apply(cbsa_race10, 1, entropyrate)
  )
cbsa_race20 <- cbsa_race20 %>% 
  mutate(
    entropy20 = apply(cbsa_race20, 1, entropyrate)
  )

```




## joining census with payscale

```{r}
# removing "micro area" and "metro area" suffix to census data to make join
# consistent for 2010
cbsa_race10$CBSA_title <- gsub(" Micro Area","", cbsa_race10$CBSA_title)
cbsa_race10$CBSA_title <- gsub(" Metro Area","", cbsa_race10$CBSA_title)
cbsa_race10$CBSA_title <- str_trim(cbsa_race10$CBSA_title) # weird but nec bc
# there are unseen prefix spaces in the cbsa titles from census data

# inner join 2010 census with payscale data
alldata <- alldata %>% inner_join(cbsa_race10, by = "CBSA_title")
cbsa_city_small %>% left_join(cbsa_race10, by = "CBSA_title")

# removing "micro area" and "metro area" suffix to census data to make join
# consistent for 2020
cbsa_race20$CBSA_title <- gsub(" Micro Area","", cbsa_race20$CBSA_title)
cbsa_race20$CBSA_title <- gsub(" Metro Area","", cbsa_race20$CBSA_title)
cbsa_race20$CBSA_title <- str_trim(cbsa_race20$CBSA_title) # weird but nec
# inner join 2020 census with payscale data
alldata <- alldata %>% inner_join(cbsa_race20, by = "CBSA_title")


# calculating entropy change over decade
alldata$d_entropy <- alldata$entropy20 - alldata$entropy10
```


## calculating changes in race
```{r}
alldata$d_white <- (alldata$t_white_2020 - alldata$t_white_2010) / alldata$t_white_2010
alldata$d_black <- (alldata$t_black_2020 - alldata$t_black_2010) / alldata$t_black_2010
alldata$d_hispanic <- (alldata$t_hispanic_2020 - alldata$t_hispanic_2010) / alldata$t_hispanic_2010
alldata$d_native <- (alldata$t_native_2020 - alldata$t_native_2010) / alldata$t_native_2010
alldata$d_asian <- (alldata$t_asian_2020 - alldata$t_asian_2010) / alldata$t_asian_2010
```

## filtering by age
```{r}
alldata2 <- alldata %>% filter(AGE>=18 & AGE <= 67)
alldata <- alldata %>% filter(AGE>=24 & AGE <= 54)
```


# Descriptive statistics
```{r}
# for women
quantile(alldata$TCC[alldata$GENDER == "Female"], probs = c(.25,.5,.75), na.rm=TRUE)
min(alldata$TCC[alldata$GENDER == "Female"], na.rm=TRUE)
max(alldata$TCC[alldata$GENDER == "Female"], na.rm=TRUE)
mean(alldata$TCC[alldata$GENDER == "Female"], na.rm=TRUE)
sd(alldata$TCC[alldata$GENDER == "Female"], na.rm=TRUE)


quantile(alldata$TCC[alldata$GENDER == "Male"], probs = c(.25,.5,.75), na.rm=TRUE)
min(alldata$TCC[alldata$GENDER == "Male"], na.rm=TRUE)
max(alldata$TCC[alldata$GENDER == "Male"], na.rm=TRUE)
mean(alldata$TCC[alldata$GENDER == "Male"], na.rm=TRUE)
sd(alldata$TCC[alldata$GENDER == "Male"], na.rm=TRUE)
```

### a little codebook:

t_2010 is total population in 2010

t_hispanic_2010 is total population of hispanics 2010
t_nonhispanic_2010 is total pop of non hispanics 2010
t_single_2010 is total pop of non hispanics identifying wit)h only one race 2010
t_white_2010 is total pop of single-race identifying whites 2010
etc


## based on quantiles of percent white
```{r}
# creating percent white
alldata$perwhite10 <- alldata$t_white_2010 / alldata$t_2010
# assigning quantiles
alldata <- alldata %>%
  mutate(whitequantile = ntile(perwhite10, 4))

gender_pay_quantiles <- function(df) {
  # initialize matrix to store
  tempmatrix <- matrix(ncol = 0, nrow = 4)
  # differentiate between male and female
  male_quantiles <- df %>% 
    filter(GENDER == "Male") %>%  dplyr::select(SALARY) %>%
  quantile(probs = c(0.25, 0.5, 0.75, 1), na.rm=TRUE)
  female_quantiles <- df %>% 
    filter(GENDER == "Female") %>%  dplyr::select(SALARY) %>%
  quantile(probs = c(0.25, 0.5, 0.75, 1), na.rm=TRUE)
  tempmatrix <- cbind(tempmatrix, male_quantiles, female_quantiles)
  return(tempmatrix)
}


# create table of quantiles across all quartiles of white
gender_quantile_table <- data.frame(matrix(ncol = 0, nrow = 4))# init table
for (i in c(1:4)){
  iter_matrix <- gender_pay_quantiles(filter(alldata, whitequantile == i))
  gender_quantile_table <- cbind(gender_quantile_table, iter_matrix)
}

# run this in a throwaway r script
kable(gender_quantile_table,
      caption = "gender pay quantiles across white quantiles",
      "latex",
      booktabs = T,
      align = c("r"),
      col.names = c("m","f","m","f","m","f","m","f")) %>%
  kable_styling("striped", full_width = F,
                position = "left", font_size = 12) %>%
  add_header_above(c(" " = 1, "1st" = 2, "2nd" = 2, "3rd" = 2, "4th" = 2))

```


## based on quantiles of percent black
```{r}
# creating percent white
alldata$perblack10 <- alldata$t_black_2010 / alldata$t_2010
# assigning quantiles
alldata <- alldata %>%
  mutate(blackquantile = ntile(perblack10, 4))



# create table of quantiles across all quartiles of white
gender_quantile_table <- data.frame(matrix(ncol = 0, nrow = 4))# init table
for (i in c(1:4)){
  iter_matrix <- gender_pay_quantiles(filter(alldata, blackquantile == i))
  gender_quantile_table <- cbind(gender_quantile_table, iter_matrix)
}

# run this in a throwaway r script
kable(gender_quantile_table,
      caption = "gender pay quantiles across black quantiles",
      "latex",
      booktabs = T,
      align = c("r"),
      col.names = c("m","f","m","f","m","f","m","f")) %>%
  kable_styling("striped", full_width = F,
                position = "left", font_size = 12) %>%
  add_header_above(c(" " = 1, "1st" = 2, "2nd" = 2, "3rd" = 2, "4th" = 2))

```


## based on quantiles of percent hispanic in 2010
```{r}
# creating percent white
alldata$perhisp10 <- alldata$t_hispanic_2010 / alldata$t_2010
# assigning quantiles
alldata <- alldata %>%
  mutate(hispquantile = ntile(perhisp10, 4))



# create table of quantiles across all quartiles of white
gender_quantile_table <- data.frame(matrix(ncol = 0, nrow = 4))# init table
for (i in c(1:4)){
  iter_matrix <- gender_pay_quantiles(filter(alldata, hispquantile == i))
  gender_quantile_table <- cbind(gender_quantile_table, iter_matrix)
}

# run this in a throwaway r script
kable(gender_quantile_table,
      caption = "gender pay quantiles across hispanic quantiles",
      "latex",
      booktabs = T,
      align = c("r"),
      col.names = c("m","f","m","f","m","f","m","f")) %>%
  kable_styling("striped", full_width = F,
                position = "left", font_size = 12) %>%
  add_header_above(c(" " = 1, "1st" = 2, "2nd" = 2, "3rd" = 2, "4th" = 2))


```



# Analysis with entropy
```{r}
alldata <- filter(alldata, GENDER %in% c("Male", "Female"))
```

## factoring data
```{r}
# factor highest education, no higher education will be baseline
alldata$highest <- relevel(as.factor(alldata$highest), ref = "none")
# factor race, white will be baseline
alldata$RACE_SET <- relevel(as.factor(alldata$RACE_SET), ref = "White")
# factor gender, male will be baseline
alldata$GENDER <- relevel(as.factor(alldata$GENDER), ref = "Male")

```


## controls only
```{r}
lm1 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest + LOCATION_STATE, 
          data = alldata)
# summary(lm1)
```

## with race
```{r}
lm2 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest + RACE_SET +
            LOCATION_STATE, 
          data = alldata)
# summary(lm2)
```

## with change in entropy
```{r}
# with change in entropy
lm3 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest + 
            RACE_SET + LOCATION_STATE + d_entropy, 
          data = alldata)
# summary(lm3)
```

## entropy * gender
```{r}
# with entropy and gender
lm4 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE + GENDER * d_entropy, 
          data = alldata)
# summary(lm4)
```


# Analysis with proportion change

## Change in white
```{r}
# with entropy and gender
lm5 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE + GENDER * d_white, 
          data = alldata)
summary(lm5)
```

## Change in black
```{r}
# with entropy and gender
lm6 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE + GENDER * d_black, 
          data = alldata)
summary(lm6)
```


## Change in hispanic
```{r}
# with entropy and gender
lm7 <- lm(log(SALARY) ~ GENDER + GENDER * d_hispanic + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE , 
          data = alldata)
summary(lm7)
```

```{r}
## southwest only
alldata$sw <- alldata$LOCATION_STATE %in% 
  c("TX", "NM", "AZ", "CA")
table(alldata$sw)

lmsw <- lm(log(SALARY) ~ GENDER + GENDER * d_hispanic + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE , 
          data = alldata[alldata$sw==TRUE,])
summary(lmsw)

lmnsw <- lm(log(SALARY) ~ GENDER + GENDER * d_hispanic + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE , 
          data = alldata[alldata$sw==FALSE,])
summary(lmnsw)
```


## Change in native american
```{r}
# with entropy and gender
lm8 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE + GENDER * d_native, 
          data = alldata)
summary(lm8)
```

## Change in asian
```{r}
# with entropy and gender
lm9 <- lm(log(SALARY) ~ GENDER + GENDER * d_asian + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE , 
          data = alldata)
summary(lm9)
```


# robustness checks

## factoring data
```{r}
# factor highest education, no higher education will be baseline
alldata2$highest <- relevel(as.factor(alldata2$highest), ref = "none")
# factor race, white will be baseline
alldata2$RACE_SET <- relevel(as.factor(alldata2$RACE_SET), ref = "White")
# factor gender, male will be baseline
alldata2$GENDER <- relevel(as.factor(alldata2$GENDER), ref = "Male")

```

## Results with larger age range
```{r}
## gender * entropy
lm10 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE + GENDER * d_entropy, 
          data = alldata2)
summary(lm10)

## change in Hispanic prop
lm11 <- lm(log(SALARY) ~ GENDER + GENDER * d_hispanic + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE , 
          data = alldata2)
summary(lm11)

  ## change in White prop
  lm12 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest +
              RACE_SET + LOCATION_STATE + GENDER * d_white, 
            data = alldata2)
  summary(lm12)

# Change in Black prop
lm13 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE + GENDER * d_black, 
          data = alldata2)
summary(lm13)

## change in asian prop
lm14 <- lm(log(SALARY) ~ GENDER + GENDER * d_asian + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE , 
          data = alldata2)
summary(lm14)

## change in native american prop
lm15 <- lm(log(SALARY) ~ GENDER + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE + GENDER * d_native, 
          data = alldata2)
summary(lm15)
```


## results for metropolitan only
```{r}
lm16 <- lm(log(SALARY) ~ GENDER + GENDER * d_hispanic + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE , 
          data = filter(alldata, 
                        metro_micro_area == "Metropolitan Statistical Area"))
summary(lm16)
```


## results for micropolitan only
```{r}
lm17 <- lm(log(SALARY) ~ GENDER + GENDER * d_hispanic + AGE + YRS_EXP + highest +
            RACE_SET + LOCATION_STATE , 
          data = filter(alldata, 
                        metro_micro_area == "Micropolitan Statistical Area"))
summary(lm17)
```


## results for high school only
```{r}
lm17 <- lm(log(SALARY) ~ GENDER + GENDER * d_hispanic + AGE + YRS_EXP +
            RACE_SET + LOCATION_STATE , 
          data = filter(alldata, 
                        highest == "none")) # no higher edu
summary(lm17)
```


## results for higher ed only
```{r}
lm18 <- lm(log(SALARY) ~ GENDER + GENDER * d_hispanic + AGE + YRS_EXP +
            RACE_SET + LOCATION_STATE , 
          data = filter(alldata, 
                        highest != "none")) # only higher edu
summary(lm18)
```





\newpage
# making tables

## entropy tables
```{r results = 'asis'}
stargazer:: stargazer(
  lm1, lm2, lm3, lm4,
  omit = c("LOCATION_STATE", "GENDERMissing", "GENDEROther" , 
           "d_entropy:GENDEROther"),
  notes = "",
  notes.append = FALSE,
  notes.align = "l",
  single.row = FALSE,
  notes.label = "",
  no.space= TRUE,
  omit.stat=c("LL","ser","f"),
  covariate.labels = c("Female", "Age", "Years of Experience", 
                       "Graduate degree", "Professional Degree", 
                       "Undergrad Degree", "Native American", "Asian", 
                       "Black", "Hispanic", "Entropy"),
  dep.var.caption  = "Dependent Variable",
  dep.var.labels   = "Log Salary"
)

```

## race prop tables

```{r results = 'asis'}
stargazer:: stargazer(
  lm5, lm6, lm7, lm8, lm9,
  omit = c("LOCATION_STATE", "GENDERMissing", "GENDEROther"),
  notes = "",
  notes.append = FALSE,
  notes.align = "l",
  single.row = FALSE,
  notes.label = "",
  no.space= FALSE,
  omit.stat=c("LL","ser","f")#,
  # covariate.labels = c("Female", "Age", "Years of Experience", 
  #                      "graduate degree", "Professional Degree", 
  #                      "Undergrad Degree", "Native American", "Asian", 
  #                      "Black", "Hispanic", "d_white", "d_black", 
  #                      "d_hispanic", "d_native", "d_asian")
)
```


# Making figures
```{r}
# a marginal effect plot of the gender gap based on change in entropy
me_interaction <- interplot(lm4, var1 = "GENDER", var2 = "d_entropy",
                            facet_labs = "", rfill = "#63aaff")+ 
  # Add labels for X and Y axes
  xlab("Change in Entropy") +
  ylab("Estimated Coefficient for Female Indicator") +
  theme(text = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(-0.2,0.6,0.1)) +
  theme_classic() + 
  theme(strip.background = element_blank())
ggsave(
  "me_interaction.png",
  plot = last_plot(),
  width = 8,
  height = 6,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)

# entropy effect marginal
library(margins)
entropy_marg <- cplot(lm4, "d_entropy", what = "effect")+ 
  # Add labels for X and Y axes
  xlab("Change in Entropy") +
  ylab("Estimated Coefficient Change in Entropy") +
  theme(text = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(-0.2,0.6,0.1))
ggsave(
  "entropy_marg.png",
  plot = last_plot(),
  width = 8,
  height = 6,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)

# gender salary barplot
gender_comparisons <- ggplot(data=alldata, aes(RACE_SET, SALARY, fill = GENDER)) +           # ggplot2 barplot with mean
  geom_bar(position = "dodge",
           stat = "summary",
           fun = "mean") + 
  theme_classic() + 
  ggtitle("Figure 3: Gender Gap by Race/Ethnicity") +
  xlab("Race/Ethnicity") +
  theme(axis.text.x = element_text(size=15, angle=0),
        legend.position = c(0.75,0.95),
        legend.title = element_blank()) +
  scale_x_discrete(labels=c("White", "Native \n American", "Asian", "Black", "Hispanic")) +
  scale_fill_manual(values=c("#00356b", "#5f712d"))+
  coord_cartesian(ylim=c(0,82500)) +
  scale_y_continuous(position="left")
ggsave(
  "gender_comparisons.png",
  plot = last_plot(),
  width = 8,
  height = 6,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)

alldata$YEAR <- year(alldata$UPDATE_DATE)
race_time1 <- ggplot() +
  geom_line(aes(x=c(2010, 2020), 
                y = c(mean(cbsa_race10$t_black_2010/cbsa_race10$t_2010),
                           mean(cbsa_race20$t_black_2020/cbsa_race20$t_2020))),
            color = "#785EF0", size = 1) +
  geom_line(aes(x=c(2010, 2020), 
                y = c(mean(cbsa_race10$t_asian_2010/cbsa_race10$t_2010),
                      mean(cbsa_race20$t_asian_2020/cbsa_race20$t_2020))),
            color = "#DC267F", size = 1) +
  geom_line(aes(x=c(2010, 2020), 
                y = c(mean(cbsa_race10$t_hispanic_2010/cbsa_race10$t_2010),
                      mean(cbsa_race20$t_hispanic_2020/cbsa_race20$t_2020))),
            color = "#FE6100", size = 1) +
  geom_line(aes(x=c(2010, 2020), 
                y = c(mean(cbsa_race10$t_native_2010/cbsa_race10$t_2010),
                      mean(cbsa_race20$t_native_2020/cbsa_race20$t_2020))),
            color = "#FFB000", size = 1) +
  scale_x_continuous(breaks = c(2010, 2020)) +
  xlab("Year") +
  ylab("Proportion of Population") +
  ggtitle("I need to add a legend")
ggsave(
  "race_time1.png",
  plot = last_plot(),
  width = 8,
  height = 3,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)

race_time2 <- ggplot() +
    geom_line(aes(x=c(2010, 2020), 
                y = c(mean(cbsa_race10$t_white_2010/cbsa_race10$t_2010),
                           mean(cbsa_race20$t_white_2020/cbsa_race20$t_2020))),
            color = "#648FFF", size = 1)
ggsave(
  "race_time2.png",
  plot = last_plot(),
  width = 8,
  height = 1,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)

# gender pay gap barplot
gender_gap <- ggplot(data=alldata, aes(GENDER, SALARY, fill = GENDER)) + 
  geom_bar(position = "dodge",
           stat = "summary",
           fun = "mean", 
           show.legend = FALSE) +
  theme_classic() + 
  theme(axis.text.x = element_text(size=15),
        axis.text.y = element_blank()) +
  scale_fill_manual(values=c("#00356b", "#5f712d")) +
  coord_cartesian(ylim=c(0,82500)) +
  scale_x_discrete(labels=c("\n Male", "\n Female"))
  #coord_flip(ylim=c(0,82599))
ggsave(
  "gender_gap.png",
  plot = last_plot(),
  width = 3.25,
  height =8,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)


# coefficient plot
library(broom)
results <- tidy(lm4)

fit_cis_95 <- confint(lm4, level = 0.95) %>% 
  data.frame() %>%
  rename("conf.low_95" = "X2.5..",
         "conf.high_95" = "X97.5..")
fit_cis_90 <- confint(lm4, level = 0.90) %>% 
  data.frame() %>%
  rename("conf.low_90" = "X5..",
         "conf.high_90" = "X95..")

results <- bind_cols(results, 
                     fit_cis_95, 
                     fit_cis_90) %>%
           rename(Variable = term,
                  Coefficient = estimate,
                  SE = std.error) %>%
           filter(Variable != "(Intercept)") %>%
           filter(Variable %in% c("GENDERFemale:d_entropy", "d_entropy", "GENDERFemale"))

coef_plot <- ggplot(results, 
       aes(x = Variable, y = Coefficient)) +
        geom_hline(yintercept = 0, 
                   colour = gray(1/2), lty = 2) +
        geom_point(aes(x = Variable, 
                    y = Coefficient)) + 
        geom_linerange(aes(x = Variable, 
                     ymin = conf.low_90,
                     ymax = conf.high_90),
                   lwd = 1) +
        geom_linerange(aes(x = Variable, 
                     ymin = conf.low_95,
                     ymax = conf.high_95),
                   lwd = 1/2) + 
        ggtitle("Estimated Effects of Gender and Entropy Change") +
        scale_x_discrete(labels = c("Entropy Change", "Female", "Entropy-Female Interaction")) +
        coord_flip()
ggsave(
  "coef_plot.png",
  plot = last_plot(),
  width = 8,
  height = 5,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)


entropy_hist <- ggplot(alldata, aes(x=d_entropy)) +
  geom_histogram(fill = "#63aaff")+
  geom_vline(aes(xintercept=0, color="red"),
            linetype="dashed",
            show.legend = FALSE) +
  scale_x_continuous(breaks = seq(-.1, .4, .1)) +
  xlim(-.1,.4) +
  ylab("Number of Observations") +
  xlab("Change in Entropy for Observation")
ggsave(
  "entropy_hist.png",
  plot = last_plot(),
  width = 8,
  height = 5,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)


avg_data <- alldata %>% group_by(YEAR) %>%
  summarise(mean_C1 = mean(t_white_2010/t_2020))

ggplot(avg_data, aes(YEAR, mean_C1)) +
  geom_point()



### interesting but not needed
# avg_data <- alldata %>% group_by(YEAR) %>%
#   summarise(mean_C1 = mean(t_white_2010/t_2020))
# 
# ggplot(avg_data, aes(YEAR, mean_C1)) +
#   geom_point()
```



```{r}
mean(alldata$t_black_2010/alldata$t_2010)
mean(alldata$t_black_2020/alldata$t_2020)
mean(alldata$t_asian_2010/alldata$t_2010)
mean(alldata$t_asian_2020/alldata$t_2020)
mean(alldata$t_hispanic_2010/alldata$t_2010)
mean(alldata$t_hispanic_2020/alldata$t_2020)
mean(alldata$t_native_2010/alldata$t_2010)
mean(alldata$t_native_2020/alldata$t_2020)
mean(alldata$t_white_2010/alldata$t_2010)
mean(alldata$t_white_2020/alldata$t_2020)

```




```{r}
# entropy effect across educations

ent_edu <- ggpredict(lm4, 
          terms = c("d_entropy"),
          condition = c(GENDER = "Male"),) %>%
  plot(colors = c("#555599")) +
  labs(title = "", x ="Change in Entropy", y = "Predicted Wage") +
  scale_color_manual(values = 1) +
  theme_classic()

ggsave(
  "ent_edu.png",
  plot = last_plot(),
  width = 8,
  height = 5,
  units = "in",
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)
```















```{r}
##############
# testing
##############
grid.arrange(gender_gap, gender_comparisons,
             layout_matrix = rbind(c(1,1,2,2,2,2,2),
                                   c(1,1,2,2,2,2,2),
                                   c(1,1,2,2,2,2,2),
                                   c(1,1,2,2,2,2,2)))


mfdf <- data.frame(
  SALARY = c(
    mean(
      alldata$SALARY[alldata$GENDER == "Male"],
      na.rm = TRUE
      ),
    mean(
      alldata$SALARY[alldata$GENDER == "Female"],
      na.rm = TRUE
      )
    ),
  GENDER = c("Male", "Female"))


df <- data.frame(dose=c("Male", "Female"),
                len=c(4.2, 10))
df2 <- data.frame(dose=c("extra"), len = c(5))
ggplot() +
  geom_bar(data = df, aes(x=dose, y=len), stat="identity") +
  geom_bar(data = df2, aes(x=dose, y=len), stat="identity")






```


